An active sampling model of mental rotation

### 2014-01-20 man/cogsci-2014/cogsci-2014.tex
Start methods section

@@

>>>>

We randomly generated 20 shapes of 5 or 6 vertices (e.g., $\Xa$ in
Figure \ref{fig:stimuli}). For each shape, we computed 20 ``same'' and
20 ``flipped'' stimuli pairs, with $\theta$ spaced at $20^\prime$
increments between 0 and 360 (with 0 and 180 degree rotations repeated
twice, in order to gather an equal number of responses for each angle
between 0 and 180). ``Same'' pairs were created by rotating $\Xa$ by
$\theta$; ``flipped'' pairs first reflected $\Xa$ across the $y$-axis,
then rotated by $\theta$.

We generated 5 shapes to be used in the practice block. Across all the
practice 10 stimuli, each shape and each angle (60, 120, 180, 240, or
300) was repeated twice (once ``flipped'' and once ``same'') such that
no shape was presented at the same angle twice. Additionally, we
generated a sixth shape to be included along with the instructions,
which had both a ``flipped'' and ``same'' version, each rotated to 320
degrees.

\subsection{Participants}

We recruited \TODO{N} participants on Amazon's Mechanical Turk. Each
participant was paid \$1.00 for approximately 15 minutes of work. Out
of these participantsm \TODO{N\% (N)} were included in our
analyses. The others were excluded either because of an experimental
error (\TODO{N\% (N)}), because they had already compelted a previous
version of the experiment (\TODO{N\% (N)}), or because they failed a
comprehension check (\TODO{N\% (N)}).

\subsection{Design}

The experiment consisted of one block of 10 practice trials followed
by two blocks of 100 experiment trials. Within a block, trials were
presented in a random order.

All participants saw the same 10 practice trials as described
above. Participants in different conditions saw different experimental
trials, however. Across participants, there were a total of 800
experimental stimuli (20 shapes $\times$ 20 angles $\times$ 2
same/flipped). These stimuli were split across conditions in the
following manner. First, stimuli were split into four blocks of 200
trials. Within each block, each shape was repeated 10 times and each
rotation was repeated 10 times (5 same, 5 flipped), such that across
all blocks, each stimulus appeared exactly once. Each block was then
split in half. Each participant then completed two half-blocks (not
necessarily both from the original full block).

@@

Participants were given the following instructions while being shown
an example ``same'' pair and an example ``flipped'' pair: \textit{``On
  each trial, you will see two images. Sometimes, they show the
  \textbf{same} object. Other times, the images show \textbf{flipped}
  objects. The task is to determine whether the two images show the
  \textbf{same} object or \textbf{flipped} objects.''}

On each trial, particpants were instructed to press the `b' key to
begin and to focus on the fixation cross that appeared for \TODO{N}
milliseconds afterwards. The two images were then presented
side-by-side, each at 300px $\times$ 300px, and participants could
press `s' to indicate they thought the images depicted the same
object, or `d' to indicate they thought the images depicted flipped
objects.

While there was no time limit on response time, participants were
urged to repond as quickly as possible while still being
accurate. Specifically, we asked them to aim for at least 85\% correct
in the experimental blocks, and provided a score counter to keep them
informed of their accuracy.
